Malacology: A Programmable Storage System

### 2016-10-21 paper/paper.tex
update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

We evaluate the performance of the ZLog Sequencer implemented in several
different configurations by showing a trade-off between aggregate throughput
achieved (shown in Figure~\ref{fig:captp}), and latency distribution (shown in
Figure~\ref{fig:capcdf}).

In Figure~\ref{fig:captp} the highest throughput labeled $1C$ shows that a
single client caching the sequencer resource may obtain new sequences at a rate
of nearly 6 million positions per second. This is not surprising given that all
state is cached locally in the client. When this resource is shared between two
clients, the MDS uses a best-effort approach to granting the sequencer resource
to each client (value $2C,BE$). This sharing has a large impact on performance,
as shown by the nearly three orders of magnitude drop off in throughput.

\begin{figure}[h]
\centering
\includegraphics{figures/capdelay-quota-behavior-a.png}
\includegraphics{figures/capdelay-quota-behavior-b.png}
\includegraphics{figures/capdelay-quota-behavior-c.png}
\caption{ Each dot in these graphs is an individual request and they are spread
randomly along the \(y\) axis. The default behavior (``None") shows
unpredictable performance, ``delay" lets each client hold the lease longer, and
``quota" gives each client the lease for a certain number of operations. Delay
improves latency and fairness while quota maximizes throughput.}
\label{fig:capdelay-quota-behavior}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics{figures/caps-delay-latency.png}
\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
latency distribution shows that if the clients hold their capabilities longer,
performance decreases. Changing the delay via policy can help the sequencer and
clients strike a balance between throughput and latency.}
\label{fig:capcdf}
\end{figure}

While a best-effort approach to sharing capabilities makes sense for a file
system client, throughput drops significantly with two clients as the majority
of the time neither client has the capability as it ping-pongs between
clients. This observation reveals a tuning parameter: how long a client
maintains its capability lease. In throughput labeled $2C,<time>$ we show
aggregate throughput as we decrease the amount of time that a client holds its
lease before responding to a request for the capability.  At lease times as
low as 10ms throughput achieve a significant proportion of the locally cached
throughput.\\

\paa{meta point: I am not sure the "The Malacology Approach" thing works, because it
makes the reader think that we are contrasting with a strawman. I think that if we replace this with (eg)
``Malacology reuses ....'', ``Malacology pulls'', etc, it reads more smoothly.}
\noindent\textbf{The Malacology Approach:} Re-use the capability management
features of the MDS.  While the batching mechanism can achieve high throughput,
it is not appropriate for applications that require predictable latency. The
throughput in Figure~\ref{fig:captp} labeled $2C,ZLog$ corresponds to a
centralized sequencer implementation. While throughput is lower for two clients
(because of the synchronous network round-trip required for each request),
Figure~\ref{fig:capcdf} shows that this approach achieves more predicable
latency compared to all other approaches except for the single client locally
cached mode.\\

\paa{there should probably be a sum-up sentence here; eg Malacology exposes knobs that
allow application programmers to tune the storage system to the needs of their application.}

\noindent\textbf{Features:} delay and quota parameters control throughput vs.
latency
\paa{this should probably be a complete sentence or sentences.  also, I don't recall
seeing any discussion of quotas}

>>>>

We evaluate the feasibility of using the metadata service to implement a
seqeuencer resource that is responsible for maintaining a total ordering of
the log. Clients contact the sequencer to obtain the tail of the log and then
independently initiate I/O, thus we measure both the throughput and latency of
obtaining new tail positions which inherently bound client append performance.

The sequencer is implemented as a custom file type in which the sequencer
state (a 64-bit integer) is embedded in the inode of a file representing the
sequencer. A total ordering of the log is imposed by the re-use of the
capability service that can be used to grant exclusive access of inode state
to clients. The metadata service is responsible for maintaining exclusivity
and granting access. Figure~\ref{fig:capdelay-quota-behavior} (a) shows the
behavior of the system in which a best-effort policy is used.  The two colors
represent points in time that the clients were able to access the resource.
The best-effort policy provides interleaving a high degree of interleaving
between clients but the system spends a large portion of time re-distributing
the capability, reducing overall throughput.

In order to control the performance of the system we implement a
policy that (1) restricts the length of time that a client may maintain
exclusive access and (2) limits the number of log positions that a client may
generate without yielding to other clients waiting for access. The behavior of
these two modes is illustrated in Figures~\ref{fig:capdelay-quota-behavior}
(b) and (c), respectively.

\begin{figure}[h]
\centering
\includegraphics{figures/capdelay-quota-behavior-a.png}
\includegraphics{figures/capdelay-quota-behavior-b.png}
\includegraphics{figures/capdelay-quota-behavior-c.png}
\caption{Each dot in these graphs is an individual request and they are spread
randomly along the \(y\) axis. The default behavior (``None") shows
unpredictable performance, ``delay" lets each client hold the lease longer, and
``quota" gives each client the lease for a certain number of operations. Delay
improves latency and fairness while quota maximizes throughput.}
\label{fig:capdelay-quota-behavior}
\end{figure}

Figure~\ref{fig:captp} demonstrates a configurable trade-off between
throughput and latency. In the experiment two clients are run each with a
fixed 0.25 second maximum reservation on the capability, and we vary the size
of the log position quota running each configuration for two minutes. The
total operations per second is the combined throughput of the two clients, and
the average latency is number of microseconds required to obtain a new log
position. With a small quota more time is spent exchanging exclusive access,
while a large quota reservation allows clients to experience a much lower
latency because they uncontended access for a longer period of time.

@@

To get a better picture of latency, Figure~\ref{fig:capcdf} shows the CDF of
latency for each client in all experiment configurations. At the 99th
percentile clients accessed the sequencer is less than a millisecond. The CDF
is cropped at the 99.999th percentile due to large outliers that we believe
occur in instances in which the MDS is performing I/O while it is in the
process of re-distributing the capability to another client.

\begin{figure}[h]
\centering
\includegraphics{figures/caps-delay-latency.png}
\caption{
The latency distribution shows that if the clients hold their capabilities longer,
performance decreases. Changing the delay via policy can help the sequencer and
clients strike a balance between throughput and latency.}
\label{fig:capcdf}
\end{figure}

We have shown how exposing the internal capability management service allows
users to navigate latency and throughput tradeoffs.  Other approaches to
designing the sequencer service also exist, such as using a centralized
service in which each access is a network round-trip. In contrast to the
mechanism we explored which is appropriate for clients with bursty workloads,
it may be easier to provide predictable performance using a centralized
service, and we will be exploring in the future how this can be achieved using
the capability system.
